---
title: "Week 2: Starting with R"
author: "UZH n UU: tested by Jonas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2: default
  bookdown::pdf_document2: default
bibliography: bibliography.bib
---
# Session Information
```{r sessionInfo, echo=FALSE, include=TRUE}
pander::pander(sessionInfo())

```{r load libraries, include=FALSE}
# install readr from CRAN
# install.packages("readr")
library(readr) # make sure your R has this library loaded, if not install it (see line above)
```

# Exercises for the R-Beginner - The fancy calculator

## Exercise 2
* from https://alexd106.github.io/intro2R/exercise_2.html

### Question 2

```{r myN, echo=TRUE, include=TRUE}
myNumber <- 12.43

# do some calculations
# natural log
log(myNumber)
```

```{r lg10, echo=TRUE, include=TRUE}
# log10
log10(myNumber)
```

```{r, echo=TRUE, include=TRUE}
log2(myNumber)
```[1] 2.520113
```{r sqrt, echo=TRUE, include=TRUE}
sqrt(myNumber)
```

```{r exp, echo=TRUE, include=TRUE}
exp(myNumber)
```
### Question 3
```{r, echo=TRUE, include=TRUE}
r <- 20
area <- pi*r*r
print(area)
```
### Question 5
```{r, echo=TRUE, include=TRUE}
childw_vec <- c(69, 62, 57, 59, 59, 64, 56, 66, 67, 66)
```

### Question 6
```{r, echo=TRUE, include=TRUE}
mean_childw<- mean(childw_vec)
print(mean_childw)
```
```{r, echo=TRUE, include=TRUE}
var(childw_vec) #variation
```
```{r, echo=TRUE, include=TRUE}
sd(childw_vec) #standard deviation
```
```{r, echo=TRUE, include=TRUE}
range(childw_vec)
```
```{r, echo=TRUE, include=TRUE}
length(childw_vec)
```
```{r, echo=TRUE, include=TRUE}
first.five<-childw_vec[1:5]
first.five
```
### Question 7
```{r, echo=TRUE, include=TRUE}
childh_vec <- c(112, 102, 83, 84, 99, 90, 77, 112, 133, 112)
summary(childh_vec)
```
```{r, echo=TRUE, include=TRUE}
some_child<-childh_vec[c(2,3,9,10)]
some_child
```
```{r, echo=TRUE, include=TRUE}
shorter_child<-childh_vec[childh_vec<=99] #any child less than/equal to 99
shorter_child
```
### Question 8
```{r, echo=TRUE, include=TRUE}
childh_vec_m<-childh_vec/100 #convert from cm to m
bmi<- childw_vec/(childh_vec_m^2) #bmi=weight/height^2
print(bmi)
```
### Question 9
```{r, echo=TRUE, include=TRUE}
seq1<- 1:10 #generate sequence from 1 to 10
seq1
```

### Question 10
```{r, echo=TRUE, include=TRUE}
seq2<-seq(from=10, to=1, by=-0.5) #generate sequence from 10 to 1 by 0.5 intervals
print(seq2) #print sequence
```

## Exercise 3

### Question 4

setwd("C:/USers/doggy/OneDrive/Documents/BIO435) #set location

### Question 5
```{r, echo=TRUE, include=TRUE}
whales<- read_tsv("mywhaledata.tsv") # load data using read_tsv (only in readr)
```
```{r, echo=TRUE, include=TRUE}
whales<-read.table("mywhaledata.tsv",
                   sep="\t", #tab based
                   header=TRUE)
View(whales)
head(whales) #first 8 columns
```

### Question 7
```{r, echo=TRUE, include=TRUE}
summary(whales) #1 NA in number of whales
```

### Question 9
```{r, echo=TRUE, include=TRUE}
deepwhale<-whales[whales$depth>1200,]
head(deepwhale)
```
```{r, echo=TRUE, include=TRUE}
steepwhale<-whales[whales$gradient>200,]
head(steepwhale)
```
```{r, echo=TRUE, include=TRUE}
quietwhale<-whales[whales$water.noise=="low",]
head(quietwhale)
```
```{r, echo=TRUE, include=TRUE}
mayloudwhale<-whales[whales$water.noise=="high" & whales$month=="May",]
head(mayloudwhale)
```
```{r, echo=TRUE, include=TRUE}
octspecialwhale<-whales[whales$water.noise=="low" & whales$month== "October" & whales$gradient>132,]
head(octspecialwhale)
```
```{r, echo=TRUE, include=TRUE}
latlongwhales<-whales[whales$latitude> 60 & whales$latitude< 61 &
                      whales$longitude> -6 & whales$longitude< -4,]
head(latlongwhales)
```
```{r, echo=TRUE, include=TRUE}
nomediumwhale<-whales[whales$water.noise != "medium",]
head(nomediumwhale)
```

###Question 10
```{r, echo=TRUE, include=TRUE}
octspecialwhale<-whales[whales$water.noise=="low" & whales$month== "October" & whales$gradient>median(whales$gradient),]
head(octspecialwhale)
```

###Question 11
```{r, echo=TRUE, include=TRUE}
manydeepwhales<-whales[whales$depth>1500 &
                       whales$number.whales>mean(whales$number.whales),] #results all NAs
View(manydeepwhales)
manydeepwhales2<-whales[whales$depth>1500 &
                        whales$number.whales>mean(whales$number.whales, na.rm=TRUE),] #use na.rm to ignore NAs
View(manydeepwhales2)
```

##Exercise 4
###Question 4
```{r, echo=TRUE, include=TRUE}
squid<-read.table("squid1.txt",
                  header= TRUE)
summary(squid)
View(squid)
```
```{r, echo=TRUE, include=TRUE}
nrow(squid) #number of rows (observations)
```
```{r, echo=TRUE, include=TRUE}
ncol(squid) #number of columns (variables)
```
```{r, echo=TRUE, include=TRUE}
squid$year_factor<-factor(squid$year) #change year from integer into factor
squid$month_factor<-factor(squid$month) #month to factor
squid$maturity_stage<-factor(squid$maturity.stage) #maturity stage to factor
str(squid)
```
###Question 5
```{r, echo=TRUE, include=TRUE}
table(squid$year_factor, squid$month_factor) #see how many observations per month and year combination 
```
```{r, echo=TRUE, include=TRUE}
xtabs(~year+month, data= squid) #see observations per month and year combo and specifies labels
#1989 missing data for months 1-11, 1990 missing 2 and 6, 1991 missing 8-12
#1990 has most observations
```
```{r, echo=TRUE, include=TRUE}
squid.table<-xtabs(~year+maturity.stage+month, data= squid) #create contingency table
squid.flat.table<-ftable(squid.table) #flatten contingency table
squid.flat.table #view table
```
###Question 9
```{r, echo=TRUE, include=TRUE}
squid.DML.w<-ggplot(squid)+
  geom_point(aes(x=DML, y=weight))+
  theme_classic()+
  labs(x= "DML",
       y= "Weight")
#linear positive relationship
```
```{r, echo=TRUE, include=TRUE}
squid$weight.log<-log(squid$weight) #log transformation
squid$weight.sqrt<-sqrt(squid$weight) #square root transformation
head(squid)
```
```{r, echo=TRUE, include=TRUE}
#scatterplot of DML vs log weight
squid.DML.logw<-ggplot(squid)+
  geom_point(aes(x=DML, y=weight.log))+
  theme_classic()+
  labs(x= "DML",
       y= "Log Weight")
```
```{r, echo=TRUE, include=TRUE}
#scatterplot of DML vs sqrt weight
squid.DML.sqrtw<-ggplot(squid)+
    geom_point(aes(x=DML, y=weight.sqrt))+
    theme_classic()+
    labs(x="DML",
         y="Square root Weight")
#best linearises relationship
```
###Question 10
```{r, echo=TRUE, include=TRUE}

```
###Question 12
```{r, echo=TRUE, include=TRUE}

```
# Session Information
```{r sessionInfo, echo=FALSE, include=TRUE}
pander::pander(sessionInfo())
```
